Relationships in the extremes and their influence on competition and coexistence

### 2022-01-11 paper/Paper.Rnw
addressing Dan's comments

%***DAN: Methods paras here for G4
To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we use a laboratory system of freshwater diatoms
explored by \cite{Descamps_2005}.
\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
resource model as follows:
put model here
%***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
After using lab experiments to parameterize the model, \cite{Descamps_2005} use simulations to
show that coexistence of \emph{C. pseudostelligera} 
and \emph{F. crotonensis} occurs when temperature fluctuates periodically, 
$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, 
%***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
%***DAN: I added units. Please check.
\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contributing mechanisms, 
and here, we consider varying parameters of temperature fluctuations and further decompose the 
coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
environmental fluctuation is temperature fluctuation. We first considered the scenario where 
\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
the coexistence decomposition for \emph{F. crotonensis}. 
%put reverse in sup mat?
%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to only consider 1 invading 2 for 
%the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
%do we need to explain how we separated the asymmetries using the sims??
%***DAN: I am not sure what you mean by this question, Jasmin
As for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
$P$, a total of three plots. For each plot, we set two of the parameters constant at the 
original parameters which coexistence was experimentally found in \cite{Descamps_2005}, 
and let the remaining temperature parameter vary. 
We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully 
large for the lottery model, compared to 
the contributions of other mechanisms. 
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).

@@

%***DAN: Jasmin, some changes are needed to the figure itself. Please make the text in the legend bigger.
%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
%(that's what it says in the legend),
%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
%"A". Make the panel labels bigger, as well as the axis numbers and labels. 
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} 
is meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
%invading which, and, at the end of the 
%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
For each panel, 
two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
and mean, $\theta_0$; see Methods) are fixed at the original values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C) and 
%***DAN: I added units for a  and \theta_0 here, please check these are correct 
the IGR and components of the MCT decomposition
are then plotted against the other parameter. The original value of the varying parameter 
is indicated by the vertical dashed magenta lines. ATA contributions 
had magnitude comparable to other components, 
%***DAN: Make sure this is still true after the correct version of C is put in, and after we
%resolve the potential problem with the wiggles on the plot. Reword if not.
indicating that ATAs can contribute meaningfully
to diatom coexistence in this system. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. 
%***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
%for these plots? If I'm right about that then please reword accordingly. 
Standard errors of plotted quantities ?.
%***DAN: Don't forget to fill in the standard error stuff
%***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
%species as the invader.
}  \label{fig:fig5}
\end{figure}

\begin{figure}
\includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
\caption{The degree of ATA contribution to the IGR change with parameters of environmental variation, i.e., temperature oscillation. The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger magnitudes signifying larger ATA contributions. In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and positive IGR and negative $\Delta_i^{[EC]}/\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and negative IGR. Thus, cyan areas indicate failure to coexist and magenta areas indicate coexistence with ATA contribution increasing with color intensity. The most intense areas of magenta indicate parameter subspaces where ATA rescue occurs. For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ >1 were capped at $|1|$. } \label{fig:fig6}
\end{figure}

>>>>

%***DAN: Methods paras here for G4
To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we use a laboratory system of freshwater diatoms
explored by \cite{Descamps_2005}.
\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
resource model as follows:
put model here
%***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
After using lab experiments to parameterize the model, \cite{Descamps_2005} use simulations to
show that coexistence of \emph{C. pseudostelligera} 
and \emph{F. crotonensis} occurs when temperature fluctuates periodically, 
$\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
%***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
%***JASMIN: You're right, it should not. Corrected
with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
%***DAN: I added units. Please check.
%***JASMIN: Checked. Done. 
\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contributing mechanisms, 
and here, we consider varying parameters of temperature fluctuations and further decompose the 
coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
environmental fluctuation is temperature fluctuation. We first considered the scenario where 
\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
the coexistence decomposition for \emph{F. crotonensis}. 
%put reverse in sup mat?
%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to only consider 1 invading 2 for 
%the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
%***JASMIN: Right. Species in the lottery model only differed in mu but here they differ in they're VKQ functions.
%*** JASMIN: Note to to run the reverse invader-resident scenario and put in SUpMat!
%do we need to explain how we separated the asymmetries using the sims??
%***DAN: I am not sure what you mean by this question, Jasmin
%***JASMIN: I mean should we explain the methodology of the code we wrote that is implemented to compute the 
% ATA contribution. The functions alignranks, normcor...  (in partialSharp_fxns.R) 
As for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
$P$, a total of three plots. For each plot, we set two of the parameters constant at the 
original parameters which coexistence was experimentally found in \cite{Descamps_2005}, 
and let the remaining temperature parameter vary. 
We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully 
large for the lottery model, compared to 
the contributions of other mechanisms. 
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).

@@

%***DAN: Jasmin, some changes are needed to the figure itself. Please make the text in the legend bigger.
%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
%(that's what it says in the legend),
%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
%"A". Make the panel labels bigger, as well as the axis numbers and labels. 
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} 
is meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
%invading which, and, at the end of the 
%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
For each panel, 
two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
and mean, $\theta_0$; see Methods) are fixed at the original values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C) and 
%***DAN: I added units for a  and \theta_0 here, please check these are correct 
%***JASMIN: Done 
the IGR and components of the MCT decomposition
are then plotted against the other parameter. The original value of the varying parameter 
is indicated by the vertical dashed magenta lines. ATA contributions 
had magnitude comparable to other components, 
%***DAN: Make sure this is still true after the correct version of C is put in, and after we
%resolve the potential problem with the wiggles on the plot. Reword if not.
indicating that ATAs can contribute meaningfully
to diatom coexistence in this system. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the invader, \emph{F. crotonensis}. 
%***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
%for these plots? If I'm right about that then please reword accordingly. 
%***JASMIN: Right, Done. 
Standard errors of plotted quantities ?.
%***DAN: Don't forget to fill in the standard error stuff
%***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
%species as the invader.
An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
\emph{F. crotonensis} is the resident.
}  \label{fig:fig5}
\end{figure}

\begin{figure}
\includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
\caption{The degree of ATA contribution to the IGR change with parameters of environmental variation, i.e., temperature 
oscillation. The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger 
magnitudes signifying larger ATA contributions. In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/
\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and positive IGR and negative $\Delta_i^{[EC]}/\text{IGR}$ only 
result from positive $\Delta_i^{[EC]}$ and negative IGR. Thus, cyan areas indicate failure to coexist and magenta areas 
indicate coexistence with ATA contribution increasing with color intensity. The most intense areas of magenta indicate 
parameter subspaces where ATA rescue occurs. For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ 
>1 were capped at $|1|$. 
An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
\emph{F. crotonensis} is the resident.} \label{fig:fig6}
\end{figure}
